---
title: "HW1 Group 5"
author: "Daniil Rodionov, Hugo Aubert"
---

# Exercise 1


```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))

data("wolfcamp")

```


Geodata class is not accepted by gstat, hence we need to convert it to sf.
```{r}
wolfcamp_sf <- st_as_sf(data.frame(X1 = wolfcamp$coords[, 1], X2 = wolfcamp$coords[, 2],
                                   values = wolfcamp$data), coords = c("X1", "X2"))
```


CV function. Here we use leave-one-out CV with standardized errors.
```{r}
idw_loocv <- function(data, p) {
  n <- nrow(data)  
  errors <- numeric(n)
  
  for (i in 1:n) {
    
    train_data <- data[-i, ] 
    test_data <- data[i, , drop = FALSE]  # drop = False, so that it does not collapse into a vector/point
    
    idw_model <- gstat(
      formula = values ~ 1, #~1 because we do not have any auxiliary variables
      locations = train_data, 
      set = list(idp = p)  #sets p
    )
    
    
    idw_prediction <- predict(idw_model, newdata = test_data)
    errors[i] <- ((idw_prediction$var1.pred - test_data$values) / sd(train_data$values))^2
  }
  
  return(mean(errors))
}
```


Here we run LOOCV over a range of possible values for p and find the optimal one.
```{r message=FALSE, warning=FALSE, include=FALSE}

p_values <- seq(0.2, 10, by = 0.2)
cv_scores <- sapply(p_values, function(p) idw_loocv(wolfcamp_sf, p))
optimal_p <- p_values[which.min(cv_scores)]
optimal_cv <- min(cv_scores)

print(optimal_p)
```



The same represented on a graph
```{r echo=TRUE, fig.height=6, fig.width=8}
cv_data <- data.frame(
  p_values = p_values,  
  cv_scores = cv_scores 
)

ggplot(cv_data, aes(x = p_values, y = cv_scores)) +             
  geom_point(color = "blue", size = 2) +           
  geom_point(aes(x = optimal_p, y = optimal_cv), color = "red", size = 3) +
  labs(title = "Cross-Validation Score for Different p Values",
       x = "p",
       y = "CV") +
  theme_minimal()  
```

# Exercise2

First, we create 20x20 grid from min to max.
```{r}
coords <- st_coordinates(wolfcamp_sf)


x_range <- seq(min(coords[, 1]), max(coords[, 1]), length.out = 20)
y_range <- seq(min(coords[, 2]), max(coords[, 2]), length.out = 20)

grid <- expand.grid(X1 = x_range, X2 = y_range)

grid_sf <- st_as_sf(grid, coords = c("X1", "X2"))
```


Then we find interpolated values for grid points. For IDW the optimal p is used. For Splines default hyperparameters n, m, h are used.

```{r message=FALSE, warning=FALSE, include=FALSE}
idw_model <- gstat(
  formula = values ~ 1, 
  locations = wolfcamp_sf, 
  set = list(idp = optimal_p)  
)

idw_prediction <- predict(idw_model, newdata = grid_sf)


grid_sf$values <- idw_prediction$var1.pred
```
```{r}
mba_result <- mba.surf(
  cbind(wolfcamp$coords[, 1], wolfcamp$coords[, 2], wolfcamp$data), 
  no.X = 20,  
  no.Y = 20,
  extend = TRUE  
)
```


LOOCV for splines. Very similar to the one for IDW. The error is lower than the one for optimal p in the IDW model.


```{r message=FALSE, warning=FALSE, include=FALSE}

#another conversion. Has to be done for compatibility
wolfcamp_df <- data.frame(
  X1 = wolfcamp$coords[, 1], 
  X2 = wolfcamp$coords[, 2], 
  values = wolfcamp$data
)

bspline_loocv <- function(data) {
  n <- nrow(data)
  errors <- numeric(n)
  
  data_matrix <- as.matrix(data)  # Ensure it's a matrix
  
  for (i in 1:n) {
    train_matrix <- data_matrix[-i, ]
    test_data <- data_matrix[i, , drop = FALSE] 
    
    test_prediction <- mba.points(
      xyz = train_matrix, 
      xy.est = matrix(c(test_data[1], test_data[2]), ncol = 2)
    )
    

    predicted_value <- test_prediction$xyz.est[3]  
    
    errors[i] <- ((predicted_value - test_data[3]) / sd(train_matrix[, 3]))^2
  }
  
  return(mean(errors))
}

cv_bspline <- bspline_loocv(wolfcamp_df)

```

```{r echo=FALSE}
# Print comparison of cross-validation errors
cat("B-spline model is marginally better:\n")
cat("B-spline Cross-Validation Error: ", cv_bspline, "\n")
cat("IDW Cross-Validation Error (Optimal p = ", optimal_p, "): ", optimal_cv, "\n")
```
Graphically, interpolated surfaces look rather similar. B-splines model produces surface which looks "smoother", "less bumpy".

```{r}
#for plotting
idw_matrix <- matrix(grid_sf$values, nrow = 20, ncol = 20)

par(mfrow = c(1, 2))

persp(
  x = x_range, y = y_range, z = idw_matrix,
  theta = 50, phi = 40, col = "lightgreen",
  xlab = "X1", ylab = "X2", zlab = "Values",
  main = "IDW Interpolation"
)

persp(
  x = mba_result$xyz.est$x, y = mba_result$xyz.est$y, z = mba_result$xyz.est$z,
  theta = 50, phi = 40, col = "lightblue",
  xlab = "X1", ylab = "X2", zlab = "Values",
  main = "B-Spline Interpolation"
)

```

