---
title: "Homework 2"
output: html_document
author: "Daniil Rodionov, Hugo Aubert"
date: "2024-10-16"
---

List of the useful packages

```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(require(corrplot)))
suppressWarnings(suppressPackageStartupMessages(require(dplyr)))
```

# Question 1

## Summary of the dataset

```{r}
load("data/shrimpsfull.RData")
summary(shrimpsdata)
```
## Histogram for the differents variable of the dataset 

```{r}
par(mfrow = c(2, 2))
for (var in names(shrimpsdata)) {
  hist(shrimpsdata[[var]], 
       main = paste("Histogramme de", var), 
       xlab = var, 
       col = "lightblue", 
       border = "black")
}
```
## Boxplot for the differents variables of the dataset

```{r}
par(mfrow = c(1, 2))
for (var in names(shrimpsdata)) {
  boxplot(shrimpsdata[[var]], 
       main = paste("Boxplot de", var), 
       xlab = var, 
       col = "lightblue", 
       border = "black")
}
```

## Correlation Matrix of the different variable of the dataset

```{r}
cor_matrix <- cor(shrimpsdata)
corrplot(cor_matrix, method = "color", tl.col = "black", tl.srt = 45,
         col = colorRampPalette(c("blue", "white", "red"))(200),
         addCoef.col = "black", 
         number.cex = 0.35,
         title = "Corrélation entre 'tot' et les autres variables")
```


## Correlation Matrix for the variable tot and the other variable of the dataset

```{r}
cor_matrix <- cor(shrimpsdata)
cor_tot <- cor_matrix[,"tot", drop = FALSE]
cor_tot <- t(cor_tot)
corrplot(cor_tot, method = "color", tl.col = "black", tl.srt = 45,
         col = colorRampPalette(c("blue", "white", "red"))(200),
         addCoef.col = "black", 
         number.cex = 0.55,
         title = "Corrélation entre 'tot' et les autres variables")

```
# Question 2

## Create a sf object with the shrimpsdata set 
```{r}
shrimpsdata_sf <- st_as_sf(data.frame(X1 = shrimpsdata$X, X2 = shrimpsdata$Y,
                                   values = shrimpsdata$tot), coords = c("X1", "X2"))
```

## CV function. Here we use leave-one-out CV with standardized errors.
```{r}
idw_loocv <- function(data, p) {
  n <- nrow(data)  
  errors <- numeric(n)
  
  for (i in 1:n) {
    
    train_data <- data[-i, ] 
    test_data <- data[i, , drop = FALSE]  # drop = False, so that it does not collapse into a vector/point
    
    idw_model <- gstat(
      formula = values ~ 1, #~1 because we do not have any auxiliary variables
      locations = train_data, 
      set = list(idp = p)  #sets p
    )
    
    
    idw_prediction <- predict(idw_model, newdata = test_data)
    errors[i] <- ((idw_prediction$var1.pred - test_data$values) / sd(train_data$values))^2
  }
  
  return(mean(errors))
}
```

## Find the optimal P for IDW Model

```{r message=FALSE, warning=FALSE, include=FALSE}

p_values <- seq(0.5, 3, by = 0.5)
cv_scores <- sapply(p_values, function(p) idw_loocv(shrimpsdata_sf, p))
optimal_p <- p_values[which.min(cv_scores)]
optimal_cv <- min(cv_scores)

print(optimal_p)
```

The same represented on a graph
```{r echo=TRUE, fig.height=6, fig.width=8}
cv_data <- data.frame(
  p_values = p_values,  
  cv_scores = cv_scores 
)

ggplot(cv_data, aes(x = p_values, y = cv_scores)) +             
  geom_point(color = "blue", size = 2) +           
  geom_point(aes(x = optimal_p, y = optimal_cv), color = "red", size = 3) +
  labs(title = "Cross-Validation Score for Different p Values",
       x = "p",
       y = "CV") +
  theme_minimal()  
```
# Question 3 

## create 20x20 grid from min to max.
```{r}
coords <- st_coordinates(shrimpsdata_sf)


x_range <- seq(min(coords[, 1]), max(coords[, 1]), length.out = 20)
y_range <- seq(min(coords[, 2]), max(coords[, 2]), length.out = 20)

grid <- expand.grid(X1 = x_range, X2 = y_range)

grid_sf <- st_as_sf(grid, coords = c("X1", "X2"))
```

Then we find interpolated values for grid points. For IDW the optimal p is used. For Splines default hyperparameters n, m, h are used.

```{r message=FALSE, warning=FALSE, include=FALSE}
idw_model <- gstat(
  formula = values ~ 1, 
  locations = shrimpsdata_sf, 
  set = list(idp = optimal_p)  
)

idw_prediction <- predict(idw_model, newdata = grid_sf)


grid_sf$values <- idw_prediction$var1.pred

cat("IDW Cross-Validation Error (Optimal p = ", optimal_p, "): ", optimal_cv, "\n")
```

```{r}
#for plotting
idw_matrix <- matrix(grid_sf$values, nrow = 20, ncol = 20)

par(mfrow = c(1, 2))

persp(
  x = x_range, y = y_range, z = idw_matrix,
  theta = 50, phi = 40, col = "lightgreen",
  xlab = "X1", ylab = "X2", zlab = "Values",
  main = "IDW Interpolation"
)
```

