---
title: "Homework 3, v2"
output: html_document
author: "Daniil Rodionov, Hugo Aubert"
date: "2024-10-31"
---

```{r}
rm(list = ls())
```

List of the useful packages

```{r message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(require(corrplot)))
suppressWarnings(suppressPackageStartupMessages(require(dplyr)))
suppressWarnings(suppressPackageStartupMessages(require(sp)))
suppressWarnings(suppressPackageStartupMessages(require(RColorBrewer)))
```

```{r warning=FALSE}
data("wolfcamp")
plot(wolfcamp)
summary(wolfcamp)
```
# Question 1

```{r warning=FALSE}
directions_deg <- c(0, 45, 90, 135)

#we need radians for variog
directions_rad <- directions_deg * (pi / 180)

variograms <- list()

for (dir in directions_rad) {
  variograms[[paste("Direction", dir * (180 / pi))]] <- variog(wolfcamp, direction = dir)
}

plot(variograms[[1]], type = "b", main = "Directional Variograms", col = "red",
     ylim = c(0, max(sapply(variograms, function(x) max(x$v)))))
for (i in 2:length(directions_rad)) {
  lines(variograms[[i]], type = "b", col = i)
}

legend("topleft", legend = paste("Direction", directions_deg, "degrees"), col = 1:length(directions_rad), lty = 1, bty = "n", adj = c(0, 0.5))



```

We can observe that the Variogram shows a linear relation between the Semi-Variance and the Distance until a distance = 250 and then goes up instead of reaching maximum (or approaching it). Also, variograms are different in different directions. Therefore, we need to use trend = 1st. 

```{r message=FALSE, warning=FALSE}
variogram_result <- variog(wolfcamp, trend = "1st")

plot(variogram_result$u, variogram_result$n, type = "b",
     main = "Number of Pairwise Comparisons per Bin",
     xlab = "distance", ylab = "count",
     col = "blue", pch = 16, lty = 1)

grid()

```





```{r warning=FALSE}
vv <- variog(wolfcamp,trend = "1st", max.dist = 250)
plot(vv,type="b")
```

Sadly, eyefit only works for .R not .Rmd files. Hence, its commented out here
```{r warning=FALSE}
#eyefit(vv) 
```


```{r warning=FALSE}
plot(vv,type="b")

vv.exp <- likfit(wolfcamp,trend = "1st", cov.model = "exponential",ini.cov.pars = c(2000,100))
lines(vv.exp, col = "red")

vv.sph <- likfit(wolfcamp, trend = "1st", cov.model = "spherical", ini.cov.pars = c(2000,100))
lines(vv.sph, col = "blue")

vv.matern <- likfit(geodata = wolfcamp, trend = "1st", cov.model = "matern", ini.cov.pars = c(2000,100), kappa = 1.2)
lines(vv.matern, col = "green")

vv.matern_2 <- likfit(geodata = wolfcamp, trend = "1st", cov.model = "matern", ini.cov.pars = c(2000,100), kappa = 0.2)
lines(vv.matern_2, col = "orange")

legend("bottomright", legend = c("Exponential", "Spherical", "Matern (1.2)", "Matern (0.2)"), 
       col = c("red", "blue", "green", "orange"), lty = 1, bty = "n")
```
```{r warning=FALSE}
# Spherical
vv.sph
summary(vv.sph)

# Exponential
vv.exp
summary(vv.exp)

# Matern (1.2)
vv.matern
summary(vv.matern)

# Matern (0.2)
vv.matern_2
summary(vv.matern_2)

```


```{r warning=FALSE}

summary_exp <- summary(vv.exp)
summary_sph <- summary(vv.sph)
summary_matern <- summary(vv.matern)
summary_matern_2 <- summary(vv.matern_2)


loglik_exp <- summary_exp$likelihood$log.L
aic_exp <- summary_exp$likelihood$AIC
bic_exp <- summary_exp$likelihood$BIC

loglik_sph <- summary_sph$likelihood$log.L
aic_sph <- summary_sph$likelihood$AIC
bic_sph <- summary_sph$likelihood$BIC

loglik_matern <- summary_matern$likelihood$log.L
aic_matern <- summary_matern$likelihood$AIC
bic_matern <- summary_matern$likelihood$BIC

loglik_matern_2 <- summary_matern_2$likelihood$log.L
aic_matern_2 <- summary_matern_2$likelihood$AIC
bic_matern_2 <- summary_matern_2$likelihood$BIC


model_comparison <- data.frame(
  Model = c("Exponential", "Spherical", "Matern (1.2)", "Matern (0.2)"),
  LogLikelihood = c(loglik_exp, loglik_sph, loglik_matern, loglik_matern_2),
  AIC = c(aic_exp, aic_sph, aic_matern, aic_matern_2),
  BIC = c(bic_exp, bic_sph, bic_matern, bic_matern_2)
)

print(model_comparison)

```

We want to maximize LogLikelihood, minimize AIC and BIC. It looks like the spherical model is the best fit here. However, the difference is very marginal.

```{r warning=FALSE}

xvalid_exp <- xvalid(wolfcamp, model = vv.exp)
xvalid_sph <- xvalid(wolfcamp, model = vv.sph)
xvalid_matern <- xvalid(wolfcamp, model = vv.matern)
xvalid_matern_2 <- xvalid(wolfcamp, model = vv.matern_2)

cv_coefficient_exp <- mean(xvalid_exp$std.error^2)
cv_coefficient_sph <- mean(xvalid_sph$std.error^2)
cv_coefficient_matern <- mean(xvalid_matern$std.error^2)
cv_coefficient_matern_2 <- mean(xvalid_matern_2$std.error^2)

xvalid_comparison <- data.frame(
  Model = c("Exponential", "Spherical", "Matern (1.2)", "Matern (0.2)"),
  CrossValidationCoefficient = c(cv_coefficient_exp, cv_coefficient_sph, 
                                 cv_coefficient_matern, cv_coefficient_matern_2)
)

print(xvalid_comparison)


```

For CV, Matern is marginally better (we want to minimize the CV). While no model is outright wrong if we need to pick one, Matern seems like the best option. It has the best CV and that's what we decided to prioritize because it indicates a better predictive ability

# Question 2


```{r warning=FALSE}

grid <- expand.grid(
  x = seq(min(wolfcamp$coords[,1]), max(wolfcamp$coords[,1])),
  y = seq(min(wolfcamp$coords[,2]), max(wolfcamp$coords[,2]))
)

kriging_result_sph <- krige.conv(wolfcamp, 
                             locations = grid, 
                             krige = krige.control(obj.model = vv.matern_2))
image(kriging_result_sph, main = "Kriging Interpolation (Matern with kappa 0.2)")
points(wolfcamp$coords, col = "red", pch = 16)




```



```{r warning=FALSE}
predictions <- kriging_result_sph$predict
std_errors <- sqrt(kriging_result_sph$krige.var)

lower_bound <- predictions - 1.96 * std_errors
upper_bound <- predictions + 1.96 * std_errors

kriging_results <- data.frame(
  x = grid$x,
  y = grid$y,
  prediction = predictions,
  lower_ci = lower_bound,
  upper_ci = upper_bound
)

```


```{r warning=FALSE}

custom_palette <- c("lightyellow", "yellow", "orange", "red", "darkred")

ggplot(kriging_results, aes(x = x, y = y)) +
  geom_tile(aes(fill = prediction)) +
  scale_fill_gradientn(colors = custom_palette) +
  labs(title = "Kriging Predictions (Matern Model with 1st-order Trend)",
       fill = "Prediction") +
  theme_minimal()

ggplot(kriging_results, aes(x = x, y = y)) +
  geom_tile(aes(fill = lower_ci)) +
  scale_fill_gradientn(colors = custom_palette) +
  labs(title = "Lower 95% Confidence Interval",
       fill = "Lower CI") +
  theme_minimal()

ggplot(kriging_results, aes(x = x, y = y)) +
  geom_tile(aes(fill = upper_ci)) +
  scale_fill_gradientn(colors = custom_palette) +
  labs(title = "Upper 95% Confidence Interval",
       fill = "Upper CI") +
  theme_minimal()


```

