---
title: "HW4"
output: html_document
author: "Daniil Rodionov, Hugo Aubert"
date: "2024-11-09"
---

```{r}
rm(list = ls())
```

# Intro

List of the useful packages

```{r message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(require(corrplot)))
suppressWarnings(suppressPackageStartupMessages(require(dplyr)))
suppressWarnings(suppressPackageStartupMessages(require(sp)))
suppressWarnings(suppressPackageStartupMessages(require(RColorBrewer)))
suppressWarnings(suppressPackageStartupMessages(require(reshape2)))
suppressWarnings(suppressPackageStartupMessages(require(FactoMineR)))
suppressWarnings(suppressPackageStartupMessages(require(factoextra)))
suppressWarnings(suppressPackageStartupMessages(require(gridExtra)))
```


```{r}
load("C:/DS/Spatial_statistics/data/shrimpsfull.RData")
load("C:/DS/Spatial_statistics/data/AllGrids.RData")
```

Year 1997 & 2006

```{r}
shrimpsdata97 <- shrimpsdata %>%
  filter(ANNO == 1997)

shrimpsdata06 <- shrimpsdata %>%
  filter(ANNO == 2006)

shrimpsdataframe06 <- data.frame(X = shrimpsdata06$X, Y =shrimpsdata06$Y, tot = (shrimpsdata06$tot + 1), bat = shrimpsdata06$bat, tempmax3p = shrimpsdata06$temp.maxq3p, temp.maxq1 = shrimpsdata06$temp.maxq1, temp.minq4p = shrimpsdata06$temp.minq4p, temp.maxq3p = shrimpsdata06$temp.maxq3p, salinity.maxq3p = shrimpsdata06$salinity.maxq3p, salinity.maxq3 = shrimpsdata06$salinity.maxq3)

shrimpsdataframe97 <- data.frame(X = shrimpsdata97$X, Y =shrimpsdata97$Y, tot = (shrimpsdata97$tot + 1), bat = shrimpsdata97$bat, tempmax3p = shrimpsdata97$temp.maxq3p, temp.maxq1 = shrimpsdata97$temp.maxq1 , temp.minq4p = shrimpsdata97$temp.minq4p , temp.maxq3p = shrimpsdata97$temp.maxq3p , salinity.maxq3p = shrimpsdata97$salinity.maxq3p , salinity.maxq3 = shrimpsdata97$salinity.maxq3)
```


# Map 

```{r message=FALSE, warning=FALSE}
shrimpsdata97 <- shrimpsdata %>% filter(ANNO == 1997)
shrimpsdata06 <- shrimpsdata %>% filter(ANNO == 2006)

color_with_zero <- c("lightgrey", "yellow", "orange", "red", "darkred")

map_1997_with_zero <- ggplot(shrimpsdata97, aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero, limits = c(0, max(shrimpsdata97$tot)), na.value = "white") +
  labs(title = "1997 (Zeroes Visible)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

map_1997_no_zero <- ggplot(shrimpsdata97 %>% filter(tot > 0), aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero[-1], limits = c(min(shrimpsdata97$tot[shrimpsdata97$tot > 0]), max(shrimpsdata97$tot)), na.value = "white") +
  labs(title = "1997 (Without Zeroes)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

map_2006_with_zero <- ggplot(shrimpsdata06, aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero, limits = c(0, max(shrimpsdata06$tot)), na.value = "white") +
  labs(title = "2006 (Zeroes Visible)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

map_2006_no_zero <- ggplot(shrimpsdata06 %>% filter(tot > 0), aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero[-1], limits = c(min(shrimpsdata06$tot[shrimpsdata06$tot > 0]), max(shrimpsdata06$tot)), na.value = "white") +
  labs(title = "2006 (Without Zeroes)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(map_1997_with_zero, map_1997_no_zero, map_2006_with_zero, map_2006_no_zero, ncol = 2)
```

The maps above are useful to identify the maximal distance which will be used for kriging. 



# Year 1997 : 

## PCA

```{r message=FALSE, warning=FALSE}
shrimpsdata97_acp <- shrimpsdata97[, sapply(shrimpsdata97, is.numeric)]
shrimpsdata97_acp <- shrimpsdata97_acp[ , !names(shrimpsdata97_acp) %in% c("ANNO")]

acp_result <- PCA(shrimpsdata97_acp, graph = FALSE)

fviz_pca_var(acp_result, 
             col.var = ifelse(rownames(acp_result$var$coord) == "tot", "grey", "red"),
             repel = TRUE, 
             title = "Variables avec mise en évidence de la variable tot")

fviz_contrib(acp_result, choice = "var", axes = 1, top = 10, title = "Axis 1")

fviz_contrib(acp_result, choice = "var", axes = 2, top = 10, title = "Axis 2")
```

## Correlation

```{r message=FALSE, warning=FALSE}
correlation_matrix97 <- cor(shrimpsdata97)
correlation_matrix_melted97 <- melt(correlation_matrix97)
correlation_matrix_tot97 <- correlation_matrix_melted97 %>%
  filter(Var1 == "tot" & Var2 != "ANNO")

# Créer une heatmap pour la ligne de la variable "tot"
ggplot(data = correlation_matrix_tot97, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2) +  # Afficher les coefficients arrondis
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Corrélation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Corrélations avec la variable tot dans shrimpsdata97", x = "", y = "") +
  coord_fixed()
```
## Variogram

```{r message=FALSE, warning=FALSE}
coords97 <- shrimpsdataframe97 %>% select(X, Y)
shrimps_geor97 <- as.geodata(shrimpsdataframe97, 
                             coords.col = 1:2, 
                             data.col = which(names(shrimpsdataframe97) == "tot"), 
                             covar.col = which(names(shrimpsdataframe97) %in% c("bat", "tempmax3p", "temp.maxq1", "temp.minq4p", "temp.maxq3p", "salinity.maxq3p", "salinity.maxq3")))

vv <- variog(shrimps_geor97)
summary(vv)
plot(vv,type="b")
```

```{r message=FALSE, warning=FALSE}

vv1 <- variog(shrimps_geor97, trend = "1st", max.dist = 60, lambda = 0)
vv2 <- variog(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
vv3 <- variog(shrimps_geor97, trend = ~ temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
vv4 <- variog(shrimps_geor97, trend = ~ salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)


plot(vv1, type = "b", col = "blue", xlab = "Distance", ylab = "Semivariance", main = "Comparison of Variograms with Different Trends")
lines(vv2$u, vv2$v, type = "b", col = "red", pch = 2)
lines(vv3$u, vv3$v, type = "b", col = "green", pch = 3)
lines(vv4$u, vv4$v, type = "b", col = "purple", pch = 4)


legend("bottomright", legend = c("Trend: 1st order", "Trend: bat + temp + salinity", "Trend: temp + salinity", "Trend: salinity only"),
       col = c("blue", "red", "green", "purple"), pch = c(1, 2, 3, 4))
```
We pick the most complex option as it seems to reach the top and level off unlike others. It looks stationary 

```{r message=FALSE, warning=FALSE}
vv <- variog(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
plot(vv, type = "b")

vv.exp <- likfit(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, 
                 cov.model = "exponential", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.exp, col = "red")

vv.sph <- likfit(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, 
                 cov.model = "spherical", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.sph, col = "blue")

vv.gau <- likfit(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, 
                 cov.model = "gaussian", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.gau, col = "green")

vv.matern1 <- likfit(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3,
                    cov.model = "matern", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1, kappa = 0.3) 

lines.variomodel(vv.matern1, col = "purple")

vv.matern2 <- likfit(shrimps_geor97, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3,
                    cov.model = "matern", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1, kappa = 1.2) # 

lines.variomodel(vv.matern2, col = "orange")

legend("bottomright", legend = c("Exponential", "Spherical", "Gaussian", "Matern (0.3)", "Matern (1.2)"),
       col = c("red", "blue", "green", "purple", "orange"), lty = 1, bty = "n")


```
As it was dicussed in the previous homework, Gaussian and Matern (for certain kappas) have a curve at the beginning. This may contradict the perceived shape of the variogram, however if the curve is small and the overall fir is good Gaussian can still perform well. In any case, we have other models and its good to have Gaussian just for comparison.


```{r include=FALSE}
cv.exp <- xvalid(shrimps_geor97, model = vv.exp)
cv.sph <- xvalid(shrimps_geor97, model = vv.sph)
cv.gau <- xvalid(shrimps_geor97, model = vv.gau)
cv.matern1 <- xvalid(shrimps_geor97, model = vv.matern1)
cv.matern2 <- xvalid(shrimps_geor97, model = vv.matern2)

error.exp <- mean(cv.exp$std.error^2)
error.sph <- mean(cv.sph$std.error^2)
error.gau <- mean(cv.gau$std.error^2)
error.matern1 <- mean(cv.matern1$std.error^2)
error.matern2 <- mean(cv.matern2$std.error^2)

```


```{r}
summary_exp <- summary(vv.exp)
summary_sph <- summary(vv.sph)
summary_gau <- summary(vv.gau)
summary_matern1 <- summary(vv.matern1)
summary_matern2 <- summary(vv.matern2)

loglik_exp <- summary_exp$likelihood$log.L
aic_exp <- summary_exp$likelihood$AIC
bic_exp <- summary_exp$likelihood$BIC

loglik_sph <- summary_sph$likelihood$log.L
aic_sph <- summary_sph$likelihood$AIC
bic_sph <- summary_sph$likelihood$BIC

loglik_gau <- summary_gau$likelihood$log.L
aic_gau <- summary_gau$likelihood$AIC
bic_gau <- summary_gau$likelihood$BIC

loglik_matern1 <- summary_matern1$likelihood$log.L
aic_matern1 <- summary_matern1$likelihood$AIC
bic_matern1 <- summary_matern1$likelihood$BIC

loglik_matern2 <- summary_matern2$likelihood$log.L
aic_matern2 <- summary_matern2$likelihood$AIC
bic_matern2 <- summary_matern2$likelihood$BIC

cv_exp <- mean(cv.exp$std.error^2)
cv_sph <- mean(cv.sph$std.error^2)
cv_gau <- mean(cv.gau$std.error^2)
cv_matern1 <- mean(cv.matern1$std.error^2)
cv_matern2 <- mean(cv.matern2$std.error^2)

model_comparison <- data.frame(
  Model = c("Exponential", "Spherical", "Gaussian", "Matern (0.3)", "Matern (1.2)"),
  LogLikelihood = c(loglik_exp, loglik_sph, loglik_gau, loglik_matern1, loglik_matern2),
  AIC = c(aic_exp, aic_sph, aic_gau, aic_matern1, aic_matern2),
  BIC = c(bic_exp, bic_sph, bic_gau, bic_matern1, bic_matern2),
  CrossValidationError = c(cv_exp, cv_sph, cv_gau, cv_matern1, cv_matern2)
)

print(model_comparison)


```
The models are all pretty close. As before, CV is the main criterion and it leads us to choose Matern with kappa = 0.3



```{r message=FALSE, warning=FALSE}


kriging_result <- krige.conv(
  geodata = shrimps_geor97,               
  locations = grid_1997[, c("X", "Y")],    
  krige = krige.control(
    obj.model = vv.matern1              
  )
)

```


```{r message=FALSE, warning=FALSE}
predicted_grid <- cbind(
  grid_1997,
  Prediction = kriging_result$predict,
  Variance = kriging_result$krige.var,
  Lower_CI = kriging_result$predict - 1.96 * sqrt(kriging_result$krige.var),
  Upper_CI = kriging_result$predict + 1.96 * sqrt(kriging_result$krige.var)
)

ggplot(predicted_grid, aes(x = X, y = Y, fill = Prediction)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Prediction Map (Matern 0.3)", fill = "Predicted Value") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Lower_CI)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Lower Bound (95% CI, Matern 0.3)", fill = "Lower 95% CI") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Upper_CI)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Upper Bound (95% CI, Matern 0.3)", fill = "Upper 95% CI") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Variance)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Variance Map (Matern 0.3)", fill = "Variance") +
  theme_minimal()

```

It should be noted that biomass, of course, cannot be negative. So, techically we could set all negative values for the lower bound to 0, but then the whole grid would look the same (0s everywhere). This issue is most likely due to large variance.



# Year 2006 :

## PCA

```{r message=FALSE, warning=FALSE}
shrimpsdata06_acp <- shrimpsdata06 %>%
    select(where(is.numeric)) %>%
    select(-matches("ANNO"))  # Exclude the year column if present

acp_result_2006 <- PCA(shrimpsdata06_acp, graph = FALSE)

fviz_pca_var(acp_result_2006,
    col.var = ifelse(rownames(acp_result_2006$var$coord) == "tot", "grey", "red"),
    repel = TRUE,
    title = "Variables with emphasis on 'tot' for 2006"
)

fviz_contrib(acp_result_2006, choice = "var", axes = 1, top = 10, title = "Axis 1 - 2006")
fviz_contrib(acp_result_2006, choice = "var", axes = 2, top = 10, title = "Axis 2 - 2006")

```

## Correlation

```{r message=FALSE, warning=FALSE}
correlation_matrix06 <- cor(shrimpsdata06)
correlation_matrix_melted06 <- melt(correlation_matrix06)
correlation_matrix_tot06 <- correlation_matrix_melted06 %>%
    filter(Var1 == "tot" & Var2 != "ANNO")

ggplot(data = correlation_matrix_tot06, aes(x = Var2, y = Var1, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 2) +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation") +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(title = "Correlations with 'tot' variable in shrimpsdata06", x = "", y = "") +
    coord_fixed()


```




```{r message=FALSE, warning=FALSE}
coords06 <- shrimpsdataframe06 %>% select(X, Y)
shrimps_geor06 <- as.geodata(shrimpsdataframe06, 
                             coords.col = 1:2, 
                             data.col = which(names(shrimpsdataframe06) == "tot"), 
                             covar.col = which(names(shrimpsdataframe06) %in% c("bat", "tempmax3p", "temp.maxq1", "temp.minq4p", "temp.maxq3p", "salinity.maxq3p", "salinity.maxq3")))

vv <- variog(shrimps_geor06)
summary(vv)
plot(vv,type="b")
```

```{r message=FALSE, warning=FALSE}
vv1 <- variog(shrimps_geor06, trend = "1st", max.dist = 60, lambda = 0)
vv2 <- variog(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
vv3 <- variog(shrimps_geor06, trend = ~ temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
vv4 <- variog(shrimps_geor06, trend = ~ salinity.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)

plot(vv1, type = "b", col = "blue", xlab = "Distance", ylab = "Semivariance", main = "Comparison of Variograms with Different Trends")
lines(vv2$u, vv2$v, type = "b", col = "red", pch = 2)
lines(vv3$u, vv3$v, type = "b", col = "green", pch = 3)
lines(vv4$u, vv4$v, type = "b", col = "purple", pch = 4)

legend("bottomright", legend = c("Trend: 1st order", "Trend: bat + temp + salinity", "Trend: temp + salinity", "Trend: salinity only"),
       col = c("blue", "red", "green", "purple"), pch = c(1, 2, 3, 4))
```


```{r message=FALSE, warning=FALSE}
vv <- variog(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3, max.dist = 60, lambda = 0)
plot(vv, type = "b")

vv.exp <- likfit(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3,
                 cov.model = "exponential", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.exp, col = "red")

vv.sph <- likfit(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3,
                 cov.model = "spherical", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.sph, col = "blue")

vv.gau <- likfit(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3,
                 cov.model = "gaussian", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1)
lines.variomodel(vv.gau, col = "green")

vv.matern1 <- likfit(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3,
                     cov.model = "matern", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1, kappa = 0.3)
lines.variomodel(vv.matern1, col = "purple")

vv.matern2 <- likfit(shrimps_geor06, trend = ~ bat + temp.maxq1 + temp.minq4p + temp.maxq3p + salinity.maxq3,
                     cov.model = "matern", ini.cov.pars = c(2000, 60), lambda = 0, nugget = 1, kappa = 1.1)
lines.variomodel(vv.matern2, col = "orange")

legend("bottomright", legend = c("Exponential", "Spherical", "Gaussian", "Matern (0.3)", "Matern (1.1)"),
       col = c("red", "blue", "green", "purple", "orange"), lty = 1, bty = "n")

```
Here the variogram does not really look stationary. It is not terribly unstable but there is no cleare range and sill. Also, in this case the Gaussian shape can be justified.


```{r message=FALSE, warning=FALSE}
cv.exp <- xvalid(shrimps_geor06, model = vv.exp)
cv.sph <- xvalid(shrimps_geor06, model = vv.sph)
cv.gau <- xvalid(shrimps_geor06, model = vv.gau)
cv.matern1 <- xvalid(shrimps_geor06, model = vv.matern1)
cv.matern2 <- xvalid(shrimps_geor06, model = vv.matern2)

error.exp <- mean(cv.exp$std.error^2)
error.sph <- mean(cv.sph$std.error^2)
error.gau <- mean(cv.gau$std.error^2)
error.matern1 <- mean(cv.matern1$std.error^2)
error.matern2 <- mean(cv.matern2$std.error^2)

summary_exp <- summary(vv.exp)
summary_sph <- summary(vv.sph)
summary_gau <- summary(vv.gau)
summary_matern1 <- summary(vv.matern1)
summary_matern2 <- summary(vv.matern2)

loglik_exp <- summary_exp$likelihood$log.L
aic_exp <- summary_exp$likelihood$AIC
bic_exp <- summary_exp$likelihood$BIC

loglik_sph <- summary_sph$likelihood$log.L
aic_sph <- summary_sph$likelihood$AIC
bic_sph <- summary_sph$likelihood$BIC

loglik_gau <- summary_gau$likelihood$log.L
aic_gau <- summary_gau$likelihood$AIC
bic_gau <- summary_gau$likelihood$BIC

loglik_matern1 <- summary_matern1$likelihood$log.L
aic_matern1 <- summary_matern1$likelihood$AIC
bic_matern1 <- summary_matern1$likelihood$BIC

loglik_matern2 <- summary_matern2$likelihood$log.L
aic_matern2 <- summary_matern2$likelihood$AIC
bic_matern2 <- summary_matern2$likelihood$BIC

model_comparison <- data.frame(
  Model = c("Exponential", "Spherical", "Gaussian", "Matern (0.3)", "Matern (1.1)"),
  LogLikelihood = c(loglik_exp, loglik_sph, loglik_gau, loglik_matern1, loglik_matern2),
  AIC = c(aic_exp, aic_sph, aic_gau, aic_matern1, aic_matern2),
  BIC = c(bic_exp, bic_sph, bic_gau, bic_matern1, bic_matern2),
  CrossValidationError = c(error.exp, error.sph, error.gau, error.matern1, error.matern2)
)

print(model_comparison)


```

Here Gaussian yeild the best CV scores.


```{r include=FALSE}
kriging_result <- krige.conv(
  geodata = shrimps_geor06,          
  locations = grid_2006[, c("X", "Y")],  
  krige = krige.control(
    obj.model = vv.gau      
  )
)
```

```{r}

predicted_grid <- cbind(
  grid_2006,
  Prediction = kriging_result$predict,
  Variance = kriging_result$krige.var,
  Lower_CI = kriging_result$predict - 1.96 * sqrt(kriging_result$krige.var),
  Upper_CI = kriging_result$predict + 1.96 * sqrt(kriging_result$krige.var)
)

ggplot(predicted_grid, aes(x = X, y = Y, fill = Prediction)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Prediction Map (Gaussian, 2006)", fill = "Predicted Value") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Lower_CI)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Lower Bound (95% CI, Gaussian, 2006)", fill = "Lower 95% CI") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Upper_CI)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Upper Bound (95% CI, Gaussian, 2006)", fill = "Upper 95% CI") +
  theme_minimal()

ggplot(predicted_grid, aes(x = X, y = Y, fill = Variance)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Kriging Variance Map (Gaussian, 2006)", fill = "Variance") +
  theme_minimal()

```

The same comment about variance. If we go back to original maps and values, obtained predictions for both years don't seem off. However, the variance for both years is very large and 0 is in the confidence interval everywhere


