---
title: "Homework 4"
output: html_document
author: "Daniil Rodionov, Hugo Aubert"
---

```{r}
rm(list = ls())
```

List of the useful packages

```{r message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(require(corrplot)))
suppressWarnings(suppressPackageStartupMessages(require(dplyr)))
suppressWarnings(suppressPackageStartupMessages(require(sp)))
suppressWarnings(suppressPackageStartupMessages(require(RColorBrewer)))
```

```{r}
load("C:/DS/Spatial_statistics/data/shrimpsfull.RData")
load("C:/DS/Spatial_statistics/data/AllGrids.RData")

```

```{r}
data <- shrimpsdata[shrimpsdata$ANNO %in% c(1997, 2006), ]
data_1997 <- data[data$ANNO == 1997, ]
data_2006 <- data[data$ANNO == 2006, ]
```

```{r}
# Use geoR package for creating empirical variograms

# Prepare data for 1997 and 2006 as geoR objects
data_1997_geo <- as.geodata(data_1997, coords.col = c("X", "Y"), data.col = "tot")
data_2006_geo <- as.geodata(data_2006, coords.col = c("X", "Y"), data.col = "tot")

# Calculate empirical variograms for both datasets without max distance restriction
variog_1997 <- variog(data_1997_geo)
variog_2006 <- variog(data_2006_geo)

# Plot the empirical variograms
plot(variog_1997, main = "Empirical Variogram for 1997", xlab = "Distance", ylab = "Semivariance", type = "b")
plot(variog_2006, main = "Empirical Variogram for 2006", xlab = "Distance", ylab = "Semivariance", type = "b")

```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(gridExtra)

# Prepare data for mapping for 1997 and 2006
shrimpsdata97 <- shrimpsdata %>% filter(ANNO == 1997)
shrimpsdata06 <- shrimpsdata %>% filter(ANNO == 2006)

# Define a color gradient that includes a light color for zero values but doesn't vanish
color_with_zero <- c("lightgrey", "yellow", "orange", "red", "darkred")

# Map for 1997 - showing zeroes as lighter color
map_1997_with_zero <- ggplot(shrimpsdata97, aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero, limits = c(0, max(shrimpsdata97$tot)), na.value = "white") +
  labs(title = "Spatial Distribution of 'tot' in 1997 (Zeroes Visible)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Map for 1997 - excluding zeroes
map_1997_no_zero <- ggplot(shrimpsdata97 %>% filter(tot > 0), aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero[-1], limits = c(min(shrimpsdata97$tot[shrimpsdata97$tot > 0]), max(shrimpsdata97$tot)), na.value = "white") +
  labs(title = "Spatial Distribution of 'tot' in 1997 (Without Zeroes)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Map for 2006 - showing zeroes as lighter color
map_2006_with_zero <- ggplot(shrimpsdata06, aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero, limits = c(0, max(shrimpsdata06$tot)), na.value = "white") +
  labs(title = "Spatial Distribution of 'tot' in 2006 (Zeroes Visible)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Map for 2006 - excluding zeroes
map_2006_no_zero <- ggplot(shrimpsdata06 %>% filter(tot > 0), aes(x = X, y = Y)) +
  geom_point(aes(color = tot), size = 2) +
  scale_color_gradientn(colors = color_with_zero[-1], limits = c(min(shrimpsdata06$tot[shrimpsdata06$tot > 0]), max(shrimpsdata06$tot)), na.value = "white") +
  labs(title = "Spatial Distribution of 'tot' in 2006 (Without Zeroes)", color = "Total") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Display maps in a 2x2 grid layout
grid.arrange(map_1997_with_zero, map_1997_no_zero, map_2006_with_zero, map_2006_no_zero, ncol = 2)


```
```
```{r}
# Plot number of observations in each distance bin
barplot(variog_1997$n, names.arg = round(variog_1997$u, 1),
        xlab = "Distance Bins", ylab = "Number of Observations",
        main = "Number of Observations in Each Distance Bin (1997)",
        col = "lightblue")
```
```{r}
# Assumes geoR package is loaded for creating empirical variograms

# Define the empirical variograms for 1997 and 2006 data with max distance set to 250
variog_1997_restricted <- variog(data_1997_geo, max.dist = 270)
variog_2006_restricted <- variog(data_2006_geo, max.dist = 270)

# Plot the restricted empirical variograms
par(mfrow = c(1, 2)) # Set up the plot area for side-by-side plotting
plot(variog_1997_restricted, main = "Empirical Variogram for 1997 (Restricted to 250 units)", 
     xlab = "Distance", ylab = "Semivariance", type = "b")
plot(variog_2006_restricted, main = "Empirical Variogram for 2006 (Restricted to 250 units)", 
     xlab = "Distance", ylab = "Semivariance", type = "b")


```
```{r}
# Assumes geoR package is loaded for creating directional empirical variograms

# Define directions in degrees and convert to radians
directions_deg <- c(0, 45, 90, 135)
directions_rad <- directions_deg * (pi / 180) # Convert to radians
colors <- c("red", "blue", "green", "purple") # Colors for each direction

# Calculate empirical variograms for different directions with max distance set to 250
variograms_1997 <- list()
variograms_2006 <- list()

for (i in seq_along(directions_rad)) {
  dir <- directions_rad[i]
  variograms_1997[[i]] <- variog(data_1997_geo, max.dist = 250, direction = dir)
  variograms_2006[[i]] <- variog(data_2006_geo, max.dist = 250, direction = dir)
}

# Adjust margins to maximize plot space
par(mar = c(5, 5, 4, 5), mfrow = c(1, 1))

# Plot the first directional variogram for 1997
plot(variograms_1997[[1]]$u, variograms_1997[[1]]$v, type = "b", col = colors[1],
     main = "Directional Variograms for 1997 (Restricted to 250)", xlab = "Distance",
     ylab = "Semivariance", xlim = c(0, 250), ylim = c(0, max(sapply(variograms_1997, function(v) max(v$v, na.rm = TRUE)))),
     cex.main = 1.2, cex.lab = 0.9, cex.axis = 0.9)

# Overlay the other directional variograms for 1997
for (i in 2:length(directions_deg)) {
    lines(variograms_1997[[i]]$u, variograms_1997[[i]]$v, type = "b", col = colors[i])
}

# Place the legend slightly inside the plot area to prevent cutoff
legend("topright", x.intersp = -0.2, y.intersp = 1.5, legend = paste("Direction", directions_deg, "°"),
       col = colors, lty = 1, pch = 1, title = "Angles", bty = "n", cex = 0.75)


# Set up global plot parameters with moderate space adjustments
par(mar = c(5, 5, 4, 5) + 0.1, mfrow = c(1, 1))



# Plot the first directional variogram for 2006 with adjusted label sizes
plot(variograms_2006[[1]]$u, variograms_2006[[1]]$v, type = "b", col = colors[1],
     main = "Directional Variograms for 2006 (Restricted to 250)", xlab = "Distance",
     ylab = "Semivariance", xlim = c(0, 250), ylim = c(0, max(sapply(variograms_2006, function(v) max(v$v, na.rm = TRUE)))),
     cex.main = 1.2, cex.lab = 1, cex.axis = 0.8)

# Overlay the other directional variograms for 2006
for (i in 2:length(directions_deg)) {
    lines(variograms_2006[[i]]$u, variograms_2006[[i]]$v, type = "b", col = colors[i])
}

# Add a legend with smaller text and adjusted position
legend("topright", inset = c(-0.05, 0), legend = paste("Direction", directions_deg, "°"),
       col = colors, lty = 1, pch = 1, title = "Angles", bty = "n", cex = 0.9)



```

```{r}
old_par <- par(no.readonly = TRUE)

par(mar = c(5, 5, 5, 5))  


options(repr.plot.width = 18, repr.plot.height = 18)  


cor_matrix <- cor(data)


corrplot(cor_matrix, 
         method = "color", 
         tl.col = "black", 
         tl.cex = 0.7,   
         tl.srt = 45,    
         col = colorRampPalette(c("blue", "white", "red"))(200), 
         addCoef.col = "black",
         number.cex = 0.3,        
         title = "Correlation between 'tot' and other variables",
         mar = c(0, 0, 2, 0))     

par(old_par)

```
```{r}
cor_tot <- cor_matrix[, "tot", drop = FALSE]
cor_tot <- t(cor_tot)  


corrplot(cor_tot, method = "color", tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200),  
         addCoef.col = "black", 
         number.cex = 0.55,       
         title = "Correlation between 'tot' and other variables",
         mar = c(0, 0, 2, 0))   
```
```{r}
# Re-create the geoR object with covariates included
data_1997_geo_full <- as.geodata(data_1997, coords.col = c("X", "Y"),
                                 data.col = "tot",
                                 covariates = ~ Recruits + Spawners + salinity.min3q + salinity.max3q +
                                               temp.min3q + temp.max3q)

data_2006_geo_full <- as.geodata(data_2006, coords.col = c("X", "Y"),
                                 data.col = "tot",
                                 covariates = ~ Recruits + Spawners + salinity.min3q + salinity.max3q +
                                               temp.min3q + temp.max3q)

```

```{r}
# Calculate the restricted empirical variograms with 1st-order trend using covariates for 1997 and 2006
variog_1997_restricted <- variog(data_1997_geo_full, max.dist = 270,
                                 trend = ~ Recruits + Spawners + 
                                          salinity.minq3 + salinity.maxq3 +
                                          salinity.minq3p + salinity.maxq3p +
                                          temp.minq3 + temp.maxq3 +
                                          temp.minq3p + temp.maxq3p)

variog_2006_restricted <- variog(data_2006_geo_full, max.dist = 270,
                                 trend = ~ Recruits + Spawners + 
                                          salinity.minq3 + salinity.maxq3 +
                                          salinity.minq3p + salinity.maxq3p +
                                          temp.minq3 + temp.maxq3 +
                                          temp.minq3p + temp.maxq3p)

```


```{r}
# Plot the restricted empirical variograms side-by-side
par(mfrow = c(1, 2))  # Set up the plot area for side-by-side comparison

# Plot for 1997
plot(variog_1997_restricted, main = "Empirical Variogram for 1997 (Restricted to 270, 1st Order Trend with Covariates)",
     xlab = "Distance", ylab = "Semivariance", type = "b")

# Plot for 2006
plot(variog_2006_restricted, main = "Empirical Variogram for 2006 (Restricted to 270, 1st Order Trend with Covariates)",
     xlab = "Distance", ylab = "Semivariance", type = "b")

```


