---
title: "HW5"
output: html_document
author: "Daniil Rodionov, Hugo Aubert"
date: "2024-11-09"
---

```{r}
rm(list = ls())
```

List of the useful packages

```{r message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(require(MASS)))
suppressWarnings(suppressPackageStartupMessages(require(MBA)))
suppressWarnings(suppressPackageStartupMessages(require(gstat)))
suppressWarnings(suppressPackageStartupMessages(require(sf)))
suppressWarnings(suppressPackageStartupMessages(require(geoR)))
suppressWarnings(suppressPackageStartupMessages(require(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(require(corrplot)))
suppressWarnings(suppressPackageStartupMessages(require(dplyr)))
suppressWarnings(suppressPackageStartupMessages(require(sp)))
suppressWarnings(suppressPackageStartupMessages(require(RColorBrewer)))
suppressWarnings(suppressPackageStartupMessages(require(reshape2)))
suppressWarnings(suppressPackageStartupMessages(require(FactoMineR)))
suppressWarnings(suppressPackageStartupMessages(require(factoextra)))
suppressWarnings(suppressPackageStartupMessages(require(fields)))
```

Year 1997 & 2006 

```{r}
shrimpsdata97 <- shrimpsdata %>%
  filter(ANNO == 1997)

shrimpsdata06 <- shrimpsdata %>%
  filter(ANNO == 2006)

# Add 1 to avoid issues with zero values and then take the logarithm
shrimpsdata97$log_tot <- log(shrimpsdata97$tot + 1)

# Add 1 to avoid issues with zero values and then take the logarithm
shrimpsdata06$log_tot <- log(shrimpsdata06$tot + 1)



shrimpsdataframe06 <- data.frame(X = shrimpsdata06$X, Y =shrimpsdata06$Y, tot = (shrimpsdata06$tot + 1), log_tot = shrimpsdata06$log_tot, bat = shrimpsdata06$bat, tempmax3p = shrimpsdata06$temp.maxq3p, temp.maxq1 = shrimpsdata06$temp.maxq1, temp.minq4p = shrimpsdata06$temp.minq4p, temp.maxq3p = shrimpsdata06$temp.maxq3p, salinity.maxq3p = shrimpsdata06$salinity.maxq3p, salinity.maxq3 = shrimpsdata06$salinity.maxq3)

shrimpsdataframe97 <- data.frame(X = shrimpsdata97$X, Y =shrimpsdata97$Y, tot = (shrimpsdata97$tot + 1), log_tot = shrimpsdata97$log_tot, bat = shrimpsdata97$bat, tempmax3p = shrimpsdata97$temp.maxq3p, temp.maxq1 = shrimpsdata97$temp.maxq1 , temp.minq4p = shrimpsdata97$temp.minq4p , temp.maxq3p = shrimpsdata97$temp.maxq3p , salinity.maxq3p = shrimpsdata97$salinity.maxq3p , salinity.maxq3 = shrimpsdata97$salinity.maxq3)

geo_data06 <- as.geodata(shrimpsdataframe06, coords.col = 1:2, data.col = 3)
geo_data97 <- as.geodata(shrimpsdataframe97, coords.col = 1:2, data.col = 3)
```

For Year 1997

```{r}
set.seed(123)

message("### Bayesian Kriging for Exponential Model ###")
kriging_bayes_exp <- krige.bayes(
  geo_data97,
  model = list(cov.model = "exponential"), 
  prior = list(
    beta.prior = "flat",
    sigmasq.prior = "reciprocal",
    phi.discrete = seq(0.1, 100, length = 20)
  ),
  locations = geo_data97$coords, 
  output = list(simulations = TRUE)
)
message("Exponential model completed")

message("### Bayesian Kriging for Spherical Model ###")
kriging_bayes_sph <- krige.bayes(
  geo_data97,
  model = list(cov.model = "spherical"), 
  prior = list(
    beta.prior = "flat",
    sigmasq.prior = "reciprocal",
    phi.discrete = seq(0.1, 100, length = 20)
  ),
  locations = geo_data97$coords, 
  output = list(simulations = TRUE)
)
message("Spherical model completed")

message("### Bayesian Kriging for Matern (0.3) Model ###")
kriging_bayes_matern03 <- krige.bayes(
  geo_data97,
  model = list(cov.model = "matern", kappa = 0.3), 
  prior = list(
    beta.prior = "flat",
    sigmasq.prior = "reciprocal",
    phi.discrete = seq(0.1, 100, length = 20)
  ),
  locations = geo_data97$coords, 
  output = list(simulations = TRUE)
)
message("Matern (0.3) model completed")

message("### Bayesian Kriging for Matern (1.2) Model ###")
kriging_bayes_matern12 <- krige.bayes(
  geo_data97,
  model = list(cov.model = "matern", kappa = 1.2), 
  prior = list(
    beta.prior = "flat",
    sigmasq.prior = "reciprocal",
    phi.discrete = seq(0.1, 100, length = 20)
  ),
  locations = geo_data97$coords, 
  output = list(simulations = TRUE)
)
message("Matern (1.2) model completed")
```



```{r}
set.seed(123)

message("### Predictions for Exponential Model ###")
kriging_predictions_exp <- krige.bayes(
  geo_data97,
  locations = grid_1997[, 1:2],  
  model = kriging_bayes_exp$model,  
  prior = list(phi.discrete = seq(0.1, 100, length = 20)), 
  output = list(simulations = TRUE)
)

mean_exp <- kriging_predictions_exp$predictive$mean
variance_exp <- kriging_predictions_exp$predictive$variance

lower_bound_exp <- mean_exp - 1.96 * sqrt(variance_exp)
upper_bound_exp <- mean_exp + 1.96 * sqrt(variance_exp)

cat("Mean predictions:", head(mean_exp))
cat("Lower bound (95%):", head(lower_bound_exp))
cat("Upper bound (95%):", head(upper_bound_exp))

message("### Predictions for Spherical Model ###")
kriging_predictions_sph <- krige.bayes(
  geo_data97,
  locations = grid_1997[, 1:2], 
  model = kriging_bayes_sph$model,  
  prior = list(phi.discrete = seq(0.1, 100, length = 20)),
  output = list(simulations = TRUE)
)

mean_sph <- kriging_predictions_sph$predictive$mean
variance_sph <- kriging_predictions_sph$predictive$variance

lower_bound_sph <- mean_sph - 1.96 * sqrt(variance_sph)
upper_bound_sph <- mean_sph + 1.96 * sqrt(variance_sph)

cat("Mean predictions:", head(mean_sph))
cat("Lower bound (95%):", head(lower_bound_sph))
cat("Upper bound (95%):", head(upper_bound_sph))

message("### Predictions for Matern (kappa = 0.3) Model ###")
kriging_predictions_matern03 <- krige.bayes(
  geo_data97,
  locations = grid_1997[, 1:2],  # Utilisation explicite des deux premières colonnes
  model = kriging_bayes_matern03$model,  # Modèle ajusté Matérn (kappa = 0.3)
  prior = list(phi.discrete = seq(0.1, 100, length = 20)),  # Ajout de phi.discrete
  output = list(simulations = TRUE)
)

# Moyenne et variance des prédictions
mean_matern03 <- kriging_predictions_matern03$predictive$mean
variance_matern03 <- kriging_predictions_matern03$predictive$variance

# Calcul des intervalles de crédibilité à 95%
lower_bound_matern03 <- mean_matern03 - 1.96 * sqrt(variance_matern03)
upper_bound_matern03 <- mean_matern03 + 1.96 * sqrt(variance_matern03)

cat("Mean predictions:", head(mean_matern03))
cat("Lower bound (95%):", head(lower_bound_matern03))
cat("Upper bound (95%):", head(upper_bound_matern03))

message("### Predictions for Matern (kappa = 1.2) Model ###")
kriging_predictions_matern12 <- krige.bayes(
  geo_data97,
  locations = grid_1997[, 1:2],  # Utilisation explicite des deux premières colonnes
  model = kriging_bayes_matern12$model,  # Modèle ajusté Matérn (kappa = 1.2)
  prior = list(phi.discrete = seq(0.1, 100, length = 20)),  # Ajout de phi.discrete
  output = list(simulations = TRUE)
)


mean_matern12 <- kriging_predictions_matern12$predictive$mean
variance_matern12 <- kriging_predictions_matern12$predictive$variance


lower_bound_matern12 <- mean_matern12 - 1.96 * sqrt(variance_matern12)
upper_bound_matern12 <- mean_matern12 + 1.96 * sqrt(variance_matern12)

cat("Mean predictions:", head(mean_matern12))
cat("Lower bound (95%):", head(lower_bound_matern12))
cat("Upper bound (95%):", head(upper_bound_matern12))
```
The Bayesian Kriging results show successful convergence across exponential, spherical, and Matern (kappa = 0.3 and kappa = 1.2) models. 

The posterior distributions highlight a preference for intermediate ranges (phi ≈ 5.36), especially in exponential and spherical models.

All models identify a spatial range of approximately 22.2, indicating a consistent spatial structure in your data.

The credibility intervals are very wide, revealing significant uncertainty, particularly in areas distant from the observation points.

The Matérn model (kappa = 0.3) seems to provide more consistent predictions and slightly narrower intervals.

```{r}
results <- data.frame(
  X = grid_1997[, 1],
  Y = grid_1997[, 2],
  Estimate = mean_matern03,       
  Variance = variance_matern03,   
  Lower_CI = lower_bound_matern03, 
  Upper_CI = upper_bound_matern03 
)
print(results)
```
```{r}
ggplot(results, aes(x = X, y = Y)) +
  geom_tile(aes(fill = Estimate)) +
  geom_point(data = as.data.frame(geo_data97$coords), aes(x = X, y = Y), color = "black", size = 1) +
  scale_fill_gradient(low = "white", high = "purple") +
  labs(
    title = "Bayesian Kriging Mean Estimates",
    x = "Longitude (X)",
    y = "Latitude (Y)"
  ) +
  theme_minimal()


ggplot(results, aes(x = X, y = Y)) +
  geom_tile(aes(fill = Lower_CI)) +
  geom_point(data = as.data.frame(geo_data97$coords), aes(x = X, y = Y), color = "black", size = 1) +
  scale_fill_gradient(low = "white", high = "purple") +
  labs(
    title = "Bayesian Kriging Lower_CI",
    x = "Longitude (X)",
    y = "Latitude (Y)"
  ) +
  theme_minimal()

ggplot(results, aes(x = X, y = Y)) +
  geom_tile(aes(fill = Upper_CI)) +
  geom_point(data = as.data.frame(geo_data97$coords), aes(x = X, y = Y), color = "black", size = 1) +
  scale_fill_gradient(low = "white", high = "purple") +
  labs(
    title = "Bayesian Kriging Upper_CI",
    x = "Longitude (X)",
    y = "Latitude (Y)"
  ) +
  theme_minimal()

ggplot(results, aes(x = X, y = Y)) +
  geom_tile(aes(fill = Variance)) +
  geom_point(data = as.data.frame(geo_data97$coords), aes(x = X, y = Y), color = "black", size = 1) +
  scale_fill_gradient(low = "white", high = "purple") +
  labs(
    title = "Bayesian Kriging Variance",
    x = "Longitude (X)",
    y = "Latitude (Y)"
  ) +
  theme_minimal()
```
```{r}
rmse <- function(observed, predicted) {
  sqrt(mean((observed - predicted)^2))
}

bayesian_xvalid <- function(geo_data, model) {
  n <- nrow(geo_data$coords)
  observed <- geo_data$data
  predictions <- numeric(n)
  errors <- numeric(n)
  
  for (i in 1:n) {
    geo_data_train <- geo_data
    geo_data_train$coords <- geo_data$coords[-i, , drop = FALSE]
    geo_data_train$data <- geo_data$data[-i]
    
    model_bayes <- krige.bayes(
      geo_data_train,
      model = model,
      prior = list(
        beta.prior = "flat",
        sigmasq.prior = "reciprocal",
        phi.discrete = seq(0.1, 100, length = 20)
      ),
      locations = geo_data$coords[i, , drop = FALSE]
    )
    
    predictions[i] <- model_bayes$predictive$mean
    errors[i] <- (observed[i] - predictions[i])^2 / model_bayes$predictive$variance
  }
  
  return(mean(errors))  
}

observed <- geo_data97$data
predictions_exp <- kriging_predictions_exp$predictive$mean          
predictions_sph <- kriging_predictions_sph$predictive$mean        
predictions_matern03 <- kriging_predictions_matern03$predictive$mean 
predictions_matern12 <- kriging_predictions_matern12$predictive$mean

rmse_exp <- rmse(observed, predictions_exp)
rmse_sph <- rmse(observed, predictions_sph)
rmse_matern03 <- rmse(observed, predictions_matern03)
rmse_matern12 <- rmse(observed, predictions_matern12)

error_bayes_exp <- bayesian_xvalid(geo_data97, list(cov.model = "exponential"))
error_bayes_sph <- bayesian_xvalid(geo_data97, list(cov.model = "spherical"))
error_bayes_matern03 <- bayesian_xvalid(geo_data97, list(cov.model = "matern", kappa = 0.3))
error_bayes_matern12 <- bayesian_xvalid(geo_data97, list(cov.model = "matern", kappa = 1.2))

results <- data.frame(
  Model = c("Exponential", "Spherical", "Matern (kappa=0.3)", "Matern (kappa=1.2)"),
  RMSE = c(rmse_exp, rmse_sph, rmse_matern03, rmse_matern12),
  Error = c(error_bayes_exp, error_bayes_sph, error_bayes_matern03, error_bayes_matern12)
)

print(results)
```

In the HW4, we had :

##          Model LogLikelihood      AIC      BIC CrossValidationError
## 1  Exponential     -264.5553 547.1107 572.1981             1.036168
## 2    Spherical     -264.3297 546.6593 571.7468             1.037848
## 3     Gaussian     -264.3535 546.7070 571.7945             1.038192
## 4 Matern (0.3)     -264.6274 547.2548 572.3423             1.036191
## 5 Matern (1.1)     -264.4646 546.9291 572.0166             1.036472

It looks that the predictions are better with the frequentist models than the Bayesian model

